The effects of binarity on planet occurrence rates measured by transit surveys

### 2017-12-21 paper/main.tex
add KM comments

To generalize the procedure of Sec.~\ref{sec:simplest}, we consider an SNR 
limited transit survey in which all the primaries and single stars have 
identical properties.
For example, they might all be solar analogs, roughly similar to 
the stars that {\it Kepler} surveyed~\citep{batalha_selection_2010}.
We let the properties of the secondaries vary, and assume that there are some 
functions $L(M)$ and $R(M)$ that specify a star's luminosity and radius in 
terms of its mass.
We then allow for an arbitrary volume-limited mass ratio distribution 
$f(q)$, and also for arbitrary true rate densities for singles, primaries, 
and doubles ($\Gamma_0, \Gamma_1, \Gamma_2$).
Since the properties of the secondaries vary, we must consider not only the 
dependence of the rate densities on planet radius, but also on stellar mass.
We presume that the observers always take the properties of a binary system to 
be those of the primary.

@@

\section{More complicated transit survey models}
\label{sec:more_complicated}
The following section applies our general equation for the apparent rate 
density (Eq.~\ref{eq:general_Gamma_a}) to study binarity's effects in more 
realistic regimes.
In general, we will write the rate density for each type of star, 
$\Gamma_i(r)$, as the product of a shape function and a constant:
\begin{equation}
\Gamma_i(r) = \frac{{\rm d}\Lambda_i}{{\rm d}r} = Z_i f_i(r),
\quad {\rm for\ }i\in\{0,1,2\},
\end{equation}
where as always, $i=0$ corresponds to single stars, $i=1$ to primaries of 
binaries, and $i=2$ to secondaries of binaries.
The shape function is normalized to unity.
The $Z_i$'s are each system type's occurrence rate $\Lambda_i$, integrated 
over all planetary radii. In other words, they are number of planets per 
single, primary, or secondary star.
This section will consider the effects of varying both $f_i(r)$ and also the
relative values of the $Z_i$'s.

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/int_rate_density_vs_radius_model_3_rpu_22.5_manyZs.pdf}{figures/int_occ_rate_vs_radius_model_3_rpu_22.5_manyZs.pdf}
    \caption{
        {\it Left:} rate density and {\it right:} rate (over $0.5r_\oplus$ 
        bins) as a function of planet radius for the distribution specified by 
        Eq.~\ref{eq:model3_radius_distribution}. This model has identical 
        primaries and single stars, and varying secondaries. We use an 
        arbitrary normalization of $Z_0=0.5$ throughout.}
    \label{fig:occ_rate_model_3_log}
\end{figure}

Though the details of the radius distribution $\Gamma_i(r)$ at $r<2r_\oplus$ 
are still an active topic of research, we can consider how binarity effects
this regime under various reasonable assumptions.
For instance, we can assume that the true radius shape function is
\begin{align}
f(r)
&\propto
\left.
\begin{cases}
r^\delta & \text{for } r\geq 2r_\oplus \\
{\rm constant} & \text{for } r\leq2r_\oplus.
\end{cases}
\right.
\label{eq:model3_radius_distribution}
\end{align}
As before, all single and primary stars in this model have identical 
properties. The secondaries have masses, radii, and luminosities that vary 
between systems: $R \propto M \propto L^{\frac{1}{\alpha}}$.
Our ``nominal model'' remains the same: 
$\alpha=3.5$, $\beta=\gamma=0$, $\delta=-2.92$.

@@

Further, Fig.~\ref{fig:frac_model_3} shows that going from apparent 
radii of $2r_\oplus$ to $\approx 1.4r_\oplus$, the fraction of detected 
planets with binary companions increases by $\approx 6-12\%$, depending on the 
assumed value of $Z_2/Z_0$.
Let the fraction of detected planets from primaries at a given apparent radius 
be $F_1(r\a)$.
In the planet-less secondaries case, one can then see by analyzing 
Eqs~\ref{eq:n_det_0} and~\ref{eq:n_det_1} semi-analytically that 
$F_1(2r_\oplus)/F_1(2r_\oplus/\sqrt{2}) \approx 1.19$.
%check_companion_fractions.py, and LB's notes 2017/12/21.0
More intuitively, this ``shift'' is a feature of any radius distribution that 
declines at large radii, and flattens below some cutoff $r_c$.
Thinking only about dilution from primaries is simplest:
at apparent radii $r\a < r_c/\sqrt{2}$, relatively more planets will be 
``diluted'' into the given $r\a$, since there are more planets in 
the undiluted distribution from $r\a$ to $\sqrt{2}r\a$.
This prediction is compared against current measurements in 
Sec.~\ref{sec:discussion}.

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/int_rate_density_vs_radius_model_4_xcut_manyZs.pdf}{figures/int_occ_rate_vs_radius_model_4_xcut_manyZs.pdf}
    \caption{
        {\it Left:} rate density and {\it right:} rate (over $0.25r_\oplus$ 
        bins) as a function of planet radius in a model with a radius gap 
        (Eq.~\ref{eq:model4_radius_distribution}).
        Similar to Fig.~\protect\ref{fig:occ_rate_model_3_log}, this model has 
        fixed primaries and single stars, but varying secondaries.
    }
    \label{fig:model_4}
\end{figure}

@@

A second assumption we used is that the detection efficiency is a 
step function in SNR.
The detection efficiencies of real transit surveys are smooth functions in 
SNR; see for example the injection \& recovery simulations performed on {\it 
    Kepler}\ data by~\citet{christiansen_measuring_2016}, and the resulting 
smooth 
functional form adopted by~\citet{fulton_california-_2017}.
If we were to take a smooth detection probability in SNR, we would need to 
include a detection efficiency term in 
Eq.~\ref{eq:defn_apparent_rate_density} to inversely weight for planets 
that had low probabilities of being detected.

The simplest way to avoid the conceptual complications of accounting for 
finite detection efficiency is to raise the SNR floor, to {\it e.g.,} 
${\rm SNR} \approx 12$ in \citet{fulton_california-_2017}'s Fig 5.
To a good approximation, this allows one to discuss stars that are 
``searchable'' and ``not searchable'' for signals of an observed depth (at 
fixed orbital period).
For {\it Kepler} specifically, coupled with high resolution imaging of a
comparison sample of non-KOI host stars, this would facilitate the process 
of directly comparing the observed KOI distribution of apparent radii with the 
distributions discussed in this study.

If one insists on using a smooth detection efficiency, then life probably 
becomes more complicated.
It might then be necessary to model the detection pipeline's efficiency while 
somehow marginalizing over binarity.
Since the relative number of detections (at fixed $\delta_{\rm obs}$) coming 
from singles, primaries, and secondaries changes as a function of the 
intrinsic distribution of planet parameters (Fig.~\ref{fig:frac_model_3}), 
some complicated deconvolution might be required.

@@

\paragraph{Fraction of detected planets with binary companions vs. apparent 
radius}
A prediction of Fig.~\ref{fig:frac_model_3} is that the fraction of detected 
planets with binary companions increases by $\approx 6-12\%$ going from 
$2r_\oplus$ to $\approx 1.4r_\oplus$.
In \citet{ziegler_robo-ao_2017}'s recent summary of the Robo-AO KOI survey, 
they reported the nearby star fraction rates for planet candidates binned by 
Earths ($r\a < 1.6r_\oplus$), Neptunes ($1.6r_\oplus < r\a < 3.9 r_\oplus$), 
Saturns ($3.9r_\oplus < r\a < 9 r_\oplus$), and Jupiters ($r\a > 9r_\oplus$).
The nearby star rates and their $1\sigma$ uncertainties 
(following~\cite{burgasser_binarity_2003}'s procedure) are as
follows:
\begin{itemize}
    \item Earths: $16.3 \pm 1.0\%$, from 1480 systems.
    \item Neptunes: $13.0 \pm 0.8\%$, from 2058 systems.
    \item Saturns: $13.6 \pm 2.0\%$, from 338 systems.
    \item Jupiters: $19.0 \pm 2.8\%$, from 247 systems.
\end{itemize} 
where the absolute normalization of the rates are less than those typical for 
Sun-like stars at least in part because of Robo-AO's sensitivity 
(see~\citealt{ziegler_robo-ao_2017}'s Fig.~2).
The uptick in the companion rate for Jupiters is likely tied to a large 
astrophysical false positive rate~\citep{santerne_sophie_2012}.
Although the data do not support as large of an uptick in companion rate 
as our Fig.~\ref{fig:frac_model_3} suggests, going from Neptunes to Earths,
there is still a weak increase in the detected planet companion fraction.

@@

We then gave a general formula for the apparent rate density inferred by an 
observer ignoring binarity (Eq.~\ref{eq:general_Gamma_a}).
As input, this equation requires a volume-limited mass ratio distribution 
$f(q)$, and also the true rate densities for planets around singles, 
primaries, and secondaries.
The assumptions that make the model tractable are:
\begin{enumerate}
    \item the transit survey is SNR limited;
    \item all the primaries and single stars have identical properties;
    \item the properties of the secondaries can vary, but there are known 
    functions $L(M)$ and $R(M)$ that specify a star's luminosity and radius in 
    terms of its mass;
    \item the binarity-ignoring observers assume that binaries have the same 
    properties as the primaries.
\end{enumerate}
Allowing for these conditions, many results follow:
\begin{itemize}
%
\item Assuming a power-law planet radius distribution, and that the same 
number of planets orbit singles, primaries, and secondaries, binarity 
influences apparent planet occurrence rates around Sun-like stars at the 
$\sim$few percent level from radii $2r_\oplus \lesssim r\a \lesssim 
17r_\oplus$ (Sec.~\ref{sec:model_2}).
%
\item Assuming a broken-power law planet radius distribution, with 
\citet{howard_planet_2012}'s exponent at $r > 2r_\oplus$ and a constant 
occurrence at $r < 2r_\oplus$, there is a ``bump'' in the apparent rate 
density at $r\a < 2r_\oplus$, leading to a relative error $\delta \Gamma_0 = 
|\Gamma_0 - \Gamma\a|/\Gamma_0$ of at most $\approx 50\%$ 
(Fig.~\ref{fig:occ_rate_model_3_log}).
Although this is smaller than current systematic uncertainties on the 
occurrence rates of Earth-sized planets, this means that binarity could
eventually become an important component of the $\eta_\oplus$ error budget.
%
\item Binarity ``fills in'' gaps in the radius distribution 
(Fig.~\ref{fig:model_4}), by an amount that could affect precise measurements 
of the depth, width, and slope of \citet{fulton_california-_2017}'s radius 
gap, in the event that planets are not carefully vetted with high resolution 
imaging.
%
\item Binarity does not lead to smaller apparent HJ occurrence rates 
(Fig.~\ref{fig:gaussian_HJ}).
This assumes a Gaussian planet radius distribution, similar to that reported 
by~\citet{petigura_CKS_2017}.
\end{itemize}

>>>>

To generalize the procedure of Sec.~\ref{sec:simplest}, we consider an SNR 
limited transit survey in which the true masses of all the singles and 
primaries are known by the observer.
We let the properties of the secondaries vary, and assume that there are some 
functions $L(M)$ and $R(M)$ that specify a star's luminosity and radius in 
terms of its mass.
We then allow for a volume-limited mass ratio distribution $f(q)$, which is 
assumed to be independent of the primary's mass.
We also allow arbitrary true rate densities for singles, primaries, 
and doubles ($\Gamma_0, \Gamma_1, \Gamma_2$).
Since the properties of the secondaries vary, we must consider not only the 
dependence of the rate densities on planet radius, but also on stellar mass.
We presume that the observers always take the properties of a binary system to 
be those of the primary.

@@

\section{More complicated transit survey models}
\label{sec:more_complicated}
The following section applies our general equation for the apparent rate 
density (Eq.~\ref{eq:general_Gamma_a}) to study binarity's effects in more 
realistic regimes.
In general, we will write the rate density for each type of star, 
$\Gamma_i(r)$, as the product of a shape function and a constant:
\begin{equation}
\Gamma_i(r) = Z_i f_i(r), \quad {\rm for\ }i\in\{0,1,2\},
\end{equation}
where as always, $i=0$ corresponds to single stars, $i=1$ to primaries of 
binaries, and $i=2$ to secondaries of binaries.
The shape function is normalized to unity.
The $Z_i$'s are each system type's occurrence rate $\Lambda_i$, integrated 
over all planetary radii. In other words, they are number of planets per 
single, primary, or secondary star.
This section will consider the effects of varying both $f_i(r)$ and also the
relative values of the $Z_i$'s.

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/int_rate_density_vs_radius_model_3_rpu_22.5_manyZs.pdf}{figures/int_occ_rate_vs_radius_model_3_rpu_22.5_manyZs.pdf}
    \caption{
        {\it Left:} rate density and {\it right:} rate (over $0.5r_\oplus$ 
        bins) as a function of planet radius for the distribution specified by 
        Eq.~\ref{eq:model3_radius_distribution}. 
        This model assumes that the true properties of all the singles and 
        primaries are known to the observer, and that the volume-limited mass 
        ratios of secondaries are drawn from a uniform distribution. We use an 
        arbitrary normalization of $Z_0=0.5$ throughout.
        The rate (left) and rate density (right) are related by $\Lambda|_a^b 
        = \int_{a}^{b}\Gamma\,{\rm d}r$.
    }
    \label{fig:occ_rate_model_3_log}
\end{figure}

Though the details of the radius distribution $\Gamma_i(r)$ at $r<2r_\oplus$ 
are still an active topic of research, we can consider how binarity effects
this regime under various reasonable assumptions.
For instance, we can assume that the true radius shape function is
\begin{align}
f(r)
&\propto
\left.
\begin{cases}
r^\delta & \text{for } r\geq 2r_\oplus \\
{\rm constant} & \text{for } r\leq2r_\oplus.
\end{cases}
\right.
\label{eq:model3_radius_distribution}
\end{align}
As before, we assume that the observers know the true properties of all 
singles and primaries. The secondaries have masses, radii, and luminosities 
that vary between systems: $R \propto M \propto L^{\frac{1}{\alpha}}$.
Our ``nominal model'' remains the same: 
$\alpha=3.5$, $\beta=\gamma=0$, $\delta=-2.92$.

@@

Further, Fig.~\ref{fig:frac_model_3} shows that going from apparent 
radii of $2r_\oplus$ to $\approx 1.4r_\oplus$, the fraction of detected 
planets with binary companions increases by $\approx 6-12\%$, depending on the 
assumed value of $Z_2/Z_0$.
Let the fraction of total detected planets from primaries at a given 
apparent radius be $F_1(r\a)$, and from secondaries let it be $F_2(r\a)$.
In the planet-less secondaries case ($F_2(r\a)=0$), one can 
analyze Eqs~\ref{eq:n_det_0} and~\ref{eq:n_det_1} semi-analytically and show 
that $F_1(2r_\oplus)/F_1(2r_\oplus/\sqrt{2}) \approx 1.19$.
%check_companion_fractions.py, and LB's notes 2017/12/21.0
More intuitively, this ``shift'' is a feature of any radius distribution that 
declines at large radii, and flattens below some cutoff $r_{\rm c}$.
This is because (for dilution about primaries)
at apparent radii $r\a < r_{\rm c}/\sqrt{2}$, relatively more planets will be 
``diluted'' into the given $r\a$, since there are more planets in 
the undiluted distribution from $r\a$ to $\sqrt{2}r\a$.
This prediction is compared against current measurements in 
Sec.~\ref{sec:discussion}.

@@

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/int_rate_density_vs_radius_model_4_xcut_manyZs.pdf}{figures/int_occ_rate_vs_radius_model_4_xcut_manyZs.pdf}
    \caption{
        {\it Left:} rate density and {\it right:} rate (over $0.25r_\oplus$ 
        bins) as a function of planet radius in a model with a radius gap 
        (Eq.~\ref{eq:model4_radius_distribution}).
        Other than the intrinsic radius distribution, this model has the same 
        assumptions as Fig.~\protect\ref{fig:occ_rate_model_3_log}.
    }
    \label{fig:model_4}
\end{figure}

@@

A second assumption we used is that the detection efficiency is a 
step function in SNR.
The detection efficiencies of real transit survey pipelines are smooth 
functions in SNR; see for example the injection \& recovery simulations 
performed by~\citet{christiansen_measuring_2016} on {\it Kepler}\ data, 
and the resulting smooth functional form adopted 
by~\citet{fulton_california-_2017}.
If we were to take a smooth detection probability in SNR, we would need to 
include a detection efficiency term in 
Eq.~\ref{eq:defn_apparent_rate_density} to inversely weight for planets 
with low probabilities of being detected.

The simplest way to avoid the conceptual complications of accounting for 
finite detection efficiency is to raise the SNR floor, to {\it e.g.,} 
${\rm SNR} \approx 12$ for {\it Kepler} 
(\citealt{fulton_california-_2017}, Fig.~5).
To a good approximation, this allows one to use the boolean distinctions 
``searchable'' and ``not searchable'' for signals of an observed depth (at 
fixed orbital period).
This also makes the survey strictly SNR-limited, rather than ``fuzzily 
SNR-limited''.
Since this threshold is set in an arbitrary manner anyway\footnote{For 
example, {\it Kepler}'s threshold of ${\rm SNR}=7.1$ was set for there to be 
no more than one statistical false alarm across the full {\it Kepler} 
campaign~\citep{jenkins_tests_2002}. Equally valid would have been to insist 
on no more than $10^{-3}$ false alarms per survey, raising the detection floor.
}, this simplification would put occurrence rate studies on 
clearer conceptual 
ground, and facilitate the process of converting the observed distributions
of apparent radii into true radius distributions.

\begin{comment}
If one insists on using a smooth detection efficiency, then life probably 
becomes more complicated.
It might then be necessary to model the detection pipeline's efficiency while 
somehow marginalizing over binarity.
Since the relative number of detections (at fixed $\delta_{\rm obs}$) coming 
from singles, primaries, and secondaries changes as a function of the 
intrinsic distribution of planet parameters (Fig.~\ref{fig:frac_model_3}), 
some complicated deconvolution might be required.
\end{comment}

@@

\paragraph{Fraction of detected planets with binary companions vs. apparent 
radius}
A prediction of Fig.~\ref{fig:frac_model_3} is that the fraction of detected 
planets with binary companions increases by $\approx 6-12\%$ going from 
$2r_\oplus$ to $\approx 1.4r_\oplus$.
In \citet{ziegler_robo-ao_2017}'s recent summary of the Robo-AO KOI survey, 
they reported the nearby star fraction rates for planet candidates binned by 
Earths ($r\a < 1.6r_\oplus$), Neptunes ($1.6r_\oplus < r\a < 3.9 r_\oplus$), 
Saturns ($3.9r_\oplus < r\a < 9 r_\oplus$), and Jupiters ($r\a > 9r_\oplus$).
The nearby star rates and their $1\sigma$ uncertainties 
(following~\cite{burgasser_binarity_2003}'s procedure) are as
follows:
\begin{itemize}
    \item Earths: $16.3 \pm 1.0\%$, from 1480 systems.
    \item Neptunes: $13.0 \pm 0.8\%$, from 2058 systems.
    \item Saturns: $13.6 \pm 2.0\%$, from 338 systems.
    \item Jupiters: $19.0 \pm 2.8\%$, from 247 systems.
\end{itemize} 
where the values of the rates are less than the $\sim 45\%$ typical for 
Sun-like stars at least in part because of Robo-AO's sensitivity 
(\citealt{ziegler_robo-ao_2017}'s Fig.~2;~\citealt{raghavan_survey_2010}).
The uptick in the companion rate for Jupiters is likely tied to a large 
astrophysical false positive rate~\citep{santerne_sophie_2012}.
Going from Neptunes to Earths, the data suggest a weak increase in the 
detected planet companion fraction, perhaps of a few percent.

@@

We then gave a general formula for the apparent rate density inferred by an 
observer ignoring binarity (Eq.~\ref{eq:general_Gamma_a}).
As input, this equation requires a volume-limited mass ratio distribution 
$f(q)$, and also the true rate densities for planets around singles, 
primaries, and secondaries.
The assumptions that make the model tractable are:
\begin{enumerate}
    \item the transit survey is SNR limited;
    \item the true properties of all the singles and primaries are known to 
    the observer;
    \item the observers assume that binaries have the same properties as the 
    primaries;
    \item there are functions $L(M)$ and $R(M)$ that specify a star's 
    luminosity and radius in terms of its mass.
\end{enumerate}
Allowing for these conditions, many results follow:
\begin{itemize}
%
\item Assuming a power-law planet radius distribution, and that the same 
number of planets orbit singles, primaries, and secondaries, binarity 
influences apparent planet occurrence rates around Sun-like stars at the 
$\sim$few percent level from radii $2r_\oplus \lesssim r\a \lesssim 
17r_\oplus$ (Sec.~\ref{sec:model_2}).
%
\item Assuming a broken-power law planet radius distribution, with 
\citet{howard_planet_2012}'s exponent at $r > 2r_\oplus$ and a constant 
occurrence at $r < 2r_\oplus$, there is a ``bump'' in the apparent rate 
density at $r\a < 2r_\oplus$, leading to a relative error $\delta \Gamma_0 = 
|\Gamma_0 - \Gamma\a|/\Gamma_0$ of at most $\approx 50\%$ 
(Fig.~\ref{fig:occ_rate_model_3_log}).
Although this is smaller than current systematic uncertainties on the 
occurrence rates of Earth-sized planets, this means that binarity could
eventually become an important component of the $\eta_\oplus$ error budget.
%
\item Binarity ``fills in'' gaps in the radius distribution 
(Fig.~\ref{fig:model_4}), by an amount that could affect precise measurements 
of the depth, width, and slope of \citet{fulton_california-_2017}'s radius 
gap, in the event that planets are not carefully vetted with high resolution 
imaging.
%
\item Binarity does not lead to smaller apparent HJ occurrence rates 
(Fig.~\ref{fig:gaussian_HJ}).
This assumes a Gaussian planet radius distribution, similar to that reported 
by~\citet{petigura_CKS_2017}.
\end{itemize}
